%
%  This is ``front matter'' for the thesis.
%
%  Regarding ``References'' below:
%      KEY    MEANING
%      PU     ``A Manual for the Preparation of Graduate Theses'',
%             The Graduate School, Purdue University, 1996.
%      TCMOS  The Chicago Manual of Style, Edition 14.
%      WNNCD  Webster's Ninth New Collegiate Dictionary.
%
%  Lines marked with "%%" may need to be changed.
%

% Statement of Thesis/Dissertation Approval Page
% This page is REQUIRED.  The page should be numbered page ``ii''
% and should NOT be listed in your TABLE OF CONTENTS.
% References: PU8 ordinal pages 5 and 29.
% The web page https://engineering.purdue.edu/AAE retrieved on
% January 8, 2017 had "School of Aeronautics and Astronautics"---that
% is used instead of "Department af Aeronautics and Astronautics"
% below.
\begin{statement}
\entry{Dr.~Andrew Mugler, Chair}{Department of Physics and Astronomy}
\entry{Dr.~Bumsoo Han}{School of Mechanical Engineering}
\entry{Dr.~Ken Ritchie}{Department of Physics and Astronomy}
\entry{Dr.~Paul Muzikar}{Department of Physics and Astronomy}
\approvedby{Dr.~John Finley}{Department of Physics and Astronomy}
\end{statement}


% Dedication page is optional.
% A name and often a message in tribute to a person or cause.
% References: PU 15, WNNCD 332.
% \begin{dedication}
%   This is the dedication.
% \end{dedication}

  % Acknowledgements page is optional but most theses include
  % a brief statement of appreciation or recognition of special
  % assistance.
  % Reference: PU 16.
\begin{acknowledgments}
  First, I would like to thank my adviser Andrew Mugler for his excellent guidance and support.
  Andrew has been a wonderful advisor, his helpful and positive attitude was critical to my success. I really appreciate the countless hours he spent helping me become a better scientist.
  I would also like to thank Bumsoo Han, Paul Muzikar, and Ken Ritchie of my thesis committee for their guidance and countless helpful conversations over the years.

  My fellow group member Sean Fancher has been an invaluable partner and helped me develop some of this work. Thank you to Hye-ran Moon and Bumsoo Han for their contributions to our collaborative project on cancer cell migration. I have enjoyed being part of our interdisciplinary team and their feedback on my ongoing projects really helped improve their value and significance.

  I want to thank the members of the Mugler research group: Tommy Byrd, Sean Fancher, Xiaoling Zhai, Shivam Gupta, Mike Vennettilli, and Andrew Mugler. Their continued support through group meetings and countless other conversations guided me through my graduate studies.
  I would also like to thank Sandy Formica, Linda Paquay, and Janice Thomaz of the department for ensuring that all aspects of my graduate career went smoothly.

  Finally, I want to give a huge thank you to all my friends, my family, and my partner Elli. Without your continual encouragement and support I truly would not have been able to accomplish this.
\end{acknowledgments}

% The preface is optional.
% References: PU 16, TCMOS 1.49, WNNCD 927.
% \begin{preface}
%   This is the preface.
% \end{preface}

  % The Table of Contents is required.
  % The Table of Contents will be automatically created for you
  % using information you supply in
  %     \chapter
  %     \section
  %     \subsection
  %     \subsubsection
  % commands.
  % Reference: PU 16.
\tableofcontents

  % If your thesis has tables, a list of tables is required.
  % The List of Tables will be automatically created for you using
  % information you supply in
  %     \begin{table} ... \end{table}
  % environments.
  % Reference: PU 16.
\listoftables

  % If your thesis has figures, a list of figures is required.
  % The List of Figures will be automatically created for you using
  % information you supply in
  %     \begin{figure} ... \end{figure}
  % environments.
  % Reference: PU 16.
\listoffigures

  % List of Symbols is optional.
  % Reference: PU 17.
% \begin{symbols}
%   $m$& mass\cr
%   $v$& velocity\cr
% \end{symbols}

  % List of Abbreviations is optional.
  % Reference: PU 17.
% \begin{abbreviations}
%   abbr& abbreviation\cr
%   bcf& billion cubic feet\cr
%   BMOC& big man on campus\cr
% \end{abbreviations}

  % Nomenclature is optional.
  % Reference: PU 17.
% \begin{nomenclature}
%   Alanine& 2-Aminopropanoic acid\cr
%   Valine& 2-Amino-3-methylbutanoic acid\cr
% \end{nomenclature}

  % Glossary is optional
  % Reference: PU 17.
% \begin{glossary}
%   chick& female, usually young\cr
%   dude& male, usually young\cr
% \end{glossary}

  % Abstract is required.
\begin{abstract}
    Cell chemotaxis is crucial to many biological functions including development, wound healing, and cancer metastasis. Chemotaxis is the process in which cells migrate in response to chemical concentration gradients. Recent experiments show that cells are capable of detecting shallow gradients as small as a 1\% concentration difference,
    and multicellular groups can improve on this by an additional order of magnitude.
    Examples from morphogenesis and metastasis demonstrate collective response to gradients equivalent to a 1 molecule difference in concentration across a cell body. While the physical constraints to cell gradient sensing are well understood, how the sensory information leads to cell migration, and coherent multicellular movement in the case of collectives, remains poorly understood.
    Here we examine how extrinsic sensory noise leads to error in chemotactic performance. First, we study single cell chemotaxis and use both simulations and analytical models to place physical constraints on chemotactic performance. Next we turn our attention to collective chemotaxis. We examine how collective cell interactions can improve chemotactic performance. We develop a novel model for quantifying the physical limit to chemotactic precision for two stereotypical modes of collective chemotaxis. Finally, we conclude by examining the effects of intercellular communication on collective chemotaxis. We use simulations to test how well collectives can chemotax through very shallow gradients with the help of communication.
    By studying these computational and theoretical models of individual and collective chemotaxis, we address the gap in knowledge between chemical sensing and directed migration.
\end{abstract}
